---
title: "Assignment1"
author: "Md Adnan Karim (7002536), Ahmad Lamloum"
date: "5/18/2020"
output:
  html_document:
    df_print: paged
  pdf_document: default
  word_document: default
---
Exercise 1.1

```{r}

library("ggplot2")
library("dplyr")
library("factoextra")
pca_toy <- read.delim("A://AI/Process/pca_toy.txt")

```

(a)
It is necesssary to standardize data before principle Component Analysis because standardization of all the variables of a dataset will have same standard deviation. In this analysis, we found that PC1 has a standard deviation of 1.72. Thus, for PC1, all variables will have same weight and will be plotted against a single axis for.

```{r}

#Standardization of pca_toy dataset
pca_df <- data.frame(scale(pca_toy))
myPr <- prcomp(pca_df)

head(pca_df)

```


(b)

```{r}
#scatter plot with PC1 and PC2




pca <- cbind(pca_df, myPr$x[,1:2])



ggplot(pca, aes(PC1, PC2)) +
  stat_ellipse(geom = "polygon", col ="black", alpha = 0.5)+
  geom_point(shape = 21, col = "black")

```


(c)
"Barplot of PC1 Percentage" shows that variable "c" contributes for above 25% of PC1. "Barplot of PC2 Percentage" shows that variable "b" contributes for above 40% of total variables in PC2. It can be stated that "c" and "b" highly contributes to PC1 and PC2, respectively.
```{r}
#determining responsible variables
var <- get_pca_var(myPr)

barplot(var$contrib[,1], main = "Barplot of PC1 Percentage", xlab = "Percent", ylab = "Variables")
barplot(var$contrib[,2], main = "Barplot of PC2 Percentage", xlab = "Percent", ylab = "Variables")
```


(d)
A total of 86.9% variation has been explained by PC1 and PC2. The importance of components chart below shows that PC1 correspond to 0.7316 of the variation, followed by PC2 with 0.1366. "My PCA Graph", shows PC1 corresponds to 73.2% and PC2 corresponds to 13.7% of variations in this dataset.
```{r}
#to determine the percentage of variance 
myPr.var <- myPr$sdev^2
myPr.var.par <- round(myPr.var/sum(myPr.var)*100,1)


barplot(myPr.var.par, main = "Scree Plot", xlab = "Principal Component",
        ylab = "Percent Variation")
p=data.frame(myPr$x)

myPr.data <- data.frame(sample=rownames(p),
                        X=p[,1],
                        Y=p[,2])
ggplot(data=myPr.data, aes(x=X, y=Y, label=sample))+
  geom_text()+
  xlab(paste("PC1 - ", myPr.var.par[1], "%", sep=""))+
  ylab(paste("PC2 - ", myPr.var.par[2], "%", sep=""))+
  theme_bw()+
  ggtitle("My PCA Graph")
summary(myPr)
```
Exercise 2.1
```{r}
library("haven")
library("VIM")
ms_toy <- read.delim("A://AI/Process/ms_toy.txt")
ms1 = ms_toy
ms2 = ms1
ms3 = ms2
ms_toy[is.na(ms_toy)]=0
mean = mean(ms_toy$ctrl.1)
stdev = sd(ms_toy$ctrl.1)
mean
stdev
```
1. Mean and Standard Deviation of Raw data with imputations:
```{r}
#Mean and Standard Deviation
new <- rbind(mean, stdev)
new
```
2. Mean value generated with lower quantile 0.25 for distribution of missing value.
The mean generated is close to the lower quantile of the raw data.
```{r}
#Mean generated from lower quantile
mean_der <- qnorm(0.25, mean)
mean_der
summary(ms_toy$ctrl.1)
```
3. A new standard deviation has been generated from 0.75 fraction
of the old Standard Deviation. 

```{r}
#new standard deviation
sd_der <- qnorm(0.75, stdev)
sd_der
#value distribution based on lower quantile of the mean
ms1[is.na(ms1$ctrl.1)]<- rnorm(ms1$ctrl.1[is.na(ms1$ctrl.1)], mean_der, sd_der)
ms1$ctrl.1
```{r}
#histogram showing the frequency of values in ctrl.1
ggplot(ms1, aes(x = ctrl.1))+
  geom_histogram(color="blue", binwidth = 0.1)+
  labs(x="intensity", y="Frequency", title = "Histogram of ctrl.1+imputed value")
```

